TacticToe: Learning the Game of Tactics

### 2018-01-22 tactictoe_jv.tex
proof search: wip

\usepackage{xspace}
\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{\arg\!\max}
\DeclareMathOperator*{\argmax}{\arg\!\max}

@@

We bridge this gap by implementing our tactical prover TacticToe on top 
of the HOL4 ITP. 
TacticToe first learns from many proof scripts
the relevance of tactics in particular proof situations.
% are effective from 
% tactic-level proofs and what kind of formulas are the results of effective 
% steps, 
This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths. 
Experiments on the \holfour standard 
library show that TacticToe is able to re-prove 66.6\% of 7123 theorems in less 
than 60 seconds. 
This is to be compared with the 42.8\% proved in the same amount of time by a 
a barebone version of TacticToe.

In the same condtion the ATP Eprover proves ..., when it is run in 
auto-schedule mode with 128 lemmas predicted and 
translated by HolyHammer.
\todo{to do eprover results running}

@@

\section{Proof Search}\label{sec:proofsearch}
\todo
{
TG:
describe the prior policy, the evaluation, 
the node selection algorithm,
the node extension (+ case of solved goal) + backpropagation.
add description of end of the algorithm.
}
Despite the best efforts of the prediction algorithms, the selected tactic may 
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in which direction.
Our search algorithm is a monte carlo tree search (MCTS) algorithm that relies 
on prior evaluation and policy functions. Those priors are learned through 
direct supervised learning in the case of the policy and via the data 
accumulated during orthogonalization for the evaluation.

So, we first give a static description of the proof tree and its component.
Then we explain how we compute the evaluation and policy functions from learned 
data. And finally, we show how one MCTS step changes the proof tree data 
structures.

@@

We assume that any set is numbered and that we can access a nth member of set 
$s$ by writing $s_n$. For readability in the next figure $(G(a_j))_i$ is 
written 
$g_j^i$.

@@

\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
\draw[-to,dotted,thick] (0) to (1m);
\draw[-to,dotted,thick] (0) to (2m);
\draw[-to,dotted,thick] (0) to (3b);
\end{tikzpicture}
\end{center}
\caption{A node of a search tree and its relation with the children of one of 
its goal}
\label{fig:choice}
\end{figure}

@@

By the base case of these definition, a goal is solved if there exists a tactic 
applied to this goal that produces no goals.

%Since the support of A gives us a bijection between tactics of the tree search 
%and the node they 
%produced, any function define on tactics of the tree search can be map to a 
%function defined on nodes.

Definition of solved goals, open goals  and pending goals.
Supervised learning introduction.

@@

\subsection{Prior tactic policy}
The estimation through the distance produced by the tactic selectors are hard 
to translate into a prior tactic policy, so we rely only on the order of 
tactics to estimate the probability with which a tactic should be tried.
Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
tactics $t_{i+1},\ldots,t_500$. We make the strong assumption that $c_policy$ 
is constant across the search. We can now calculate the prior policy by:

@@

@@

For all evaluations functions, if a tactic is not productive then there is no 
output to evaluate, so the result is a failure and a reward of $0$ will be 
given for this node. The same thing happens if the node selected is already 
proven or a descendant of a proven goal. 
Experimentally, it is preferable not to 
filter proven (or descendants of proven) nodes before 
node selection to keep the estimation of the algorithm balanced.

@@

A similar trade-off between exploration and exploitation exists in algorithm 
like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
simulations and do not reflect our exploration strategy that only

We present here a monte carlo tree search with no roll out which is essentially 
the same as the one use by AlphaGo Zero~\cite{} originally defined by \cite{} 
in \cite{}.

@@

The exploration term is determined by the prior policy and the current policy 
where as the exploitation is exactly the current evaluation build from prior of 
evaluations of descendants of $a_i$.

\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]

\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The 
square root skews this probability to favor even more exploration of nodes with 
few visits.

The exploitation term is the average of the evaluation of 
all nodes descendant of $a_i$ (itself included) in the current search tree. It 
is also called the current evaluation of $a_i$.   
\[Exploitation(a_i) = CurEvaluation(a_i) = 
  Average_{b \in Desc(a_i)} {PriorEvaluation(b)}\]

MCTS with prior policy
This modification of the search tree can be simulated by weighting each arrow 
by the distance shown by the diagonalized tree. To simulated the 
diagonalization in the original tree, the arrow $t_i$ will be given weight $i$.
Methods for exploring a tree with a large branching factor has been explored  
in Go~\cite{alphago} and

However for very large branching factor RTS Combinatorial Multi-armed Bandits
for Real-Time Strategy Games. The method of choice for solving those 
problems 
is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
to each branches

In the context of changing wiehgts, the diagonalized tree has the advantage of 
being a little faster to explore as the ordered of tactic is inscribed in the 
shape of the tree. if the prior shapes

\subsection{Node extension}

@@

\begin{definition} (Productive tactic)
During a proof search, after a tactic $t$ is applied to a node with active goal 
$g$, 
$t$ is called 
productive if and only if all the following conditions are satisified:
\begin{itemize}
\item It does not fail. We note $l$ the list of goals returned by the tactic.
\item It does not loop. $l$ do not contain any goals that appears in its 
ancestor nodes.
\item It is not a parallel step. $l$ is not a superset of a list of goals 
produced from the same node with the same active goal.
\end{itemize}

@@

\subsection{End of the proof search}
The proof search stops when one of these 3 conditions is realized:
\begin{itemize}
\item It found a proof (i.e. the root node is solved). In this case, 
the search returns a minimized and 
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It has saturated. 
There is no more tactics to be applied to any open goal. \todo{define open 
goal}
\item It exceeded a given time limit.
\end{itemize}

>>>>

\usepackage{xspace}
\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{\arg\!\max}
\DeclareMathOperator*{\argmax}{\arg\!\max}
\DeclareMathOperator*{\average}{Average}

@@

We bridge this gap by implementing our tactical prover TacticToe on top 
of the HOL4 ITP. 
TacticToe first learns from many proof scripts
the relevance of tactics in particular proof situations.
% are effective from 
% tactic-level proofs and what kind of formulas are the results of effective 
% steps, 
This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths. 
Experiments on the \holfour standard 
library show that TacticToe is able to re-prove 66.2\% of all 7164 theorems in 
less 
than 60 seconds. 
This is to be compared with the 42.9\% proved in the same amount of time by 
Eprover in auto-schedule mode with 128 lemmas predicted and 
translated by HolyHammer.
\todo{run experiments including proof scripts containing let
     run experiments without prove on 1/10th of the library.}

@@

\section{Proof Search}\label{sec:proofsearch}
\todo
{
TG:
describe the prior policy, the evaluation, 
the node selection algorithm,
the node extension (+ case of solved goal) + backpropagation.
add description of end of the algorithm.
}
Despite the best efforts of the prediction algorithms, the selected tactic may 
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
on prior evaluation and policy functions. Those priors are learned through 
direct supervised learning in the case of the policy and via the data 
accumulated during orthogonalization for the evaluation. This MCTS do not rely 
on roll outs. The 
policy and evaluation are estimated by the simple supervised k-nn algorithm 
here instead of neural networks trained through reinforcement learning in 
AlphaGo Zero~\cite{}.

%However for very large branching factor RTS Combinatorial 
%Multi-armed Bandits
%for Real-Time Strategy Games. The method of choice for solving those 
%problems 
%is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
%to each branches

So, we first give a static description of the proof tree and its component.
Then we explain how we compute the evaluation and policy functions from learned 
data. And finally, we show how the two MCTS step, node selection and node 
extension changes the proof tree data structures.

@@

Unless specified otherwise, we assume that any set is numbered arbitrarily and 
that we can access a nth member of set 
$s$ by writing $s_n$. For readability in the next figure $(G(a_j))_i$ is 
written $g_j^i$.

@@

\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
\draw[-to,dotted,thick] (0) to (1m);
\draw[-to,dotted,thick] (0) to (2m);
\draw[-to,dotted,thick] (0) to (3b);
\end{tikzpicture}
\end{center}
\caption{A node of a search tree and the children of one of 
its goal.}
\label{fig:choice}
\end{figure}

@@

By the base case of these definition, a goal is solved if there exists a tactic 
applied to this goal that produces no goals.
\end{definition}

@@

\subsection{Prior tactic policy}
The estimation through the distance produced by the tactic selectors are hard 
to translate into a prior tactic policy, so we rely only on the order of 
tactics to estimate the probability with which a tactic should be tried.
Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
tactics $t_{i+1},\ldots,t_{500}$. We make the strong assumption that $c_policy$ 
is constant across the search. We can now calculate the prior policy by:

@@

Since the support of A induces a bijection between tactics of the tree 
search and the node they 
produced, any function define on tactics of the tree search can be map to a 
function defined on nodes. 
Therefore, we define $PriorPolicy(a)=_def PriorPolicy(t)$, if $\exists g 
A(g,t) = a$.

@@

For all evaluations functions, if a tactic is not productive then there is no 
output to evaluate, so the result is a failure and a reward of $0$ will be 
given for this node. The same thing happens if the node selected is already 
proven or a descendant of a proven goal. 
Experimentally, it is preferable not to 
filter solved (or descendants of solved) nodes before 
node selection to keep the estimation of the algorithm balanced.

\subsection{Start of the search}
The input of the algorithm is a goal $g$ that we want to prove.
Therefore the search tree starts with only one node containing the singleton 
$g$ and an ordered list of predicted tactics for the goal $g$.

@@

%A similar trade-off between exploration and exploitation exists in algorithm 
%like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
%simulations and do not reflect our exploration strategy that only

Through node extension steps the tree grows and  the number of paths to explore 
increases. To decide which node to extend next, we proceed through the node 
selection algorithm. Starting from the root node, we repeatedly choses the 
children with the highest value. To allow the algorithm to select nodes that 
are not leafs, We also assign a value to the possibility of stopping the 
descent through the tree at a particular node and trying the next tactic in the 
predictions. This value has an exploitation equal to 0 since it was never 
explored.

The value of a node with respect to a search tree, prior policies and prior 
evaluation is determined by the 
following computation:

@@

The exploration term is determined by the prior policy and the current policy 
where as the exploitation is exactly the current evaluation build from prior of 
evaluations of descendants of $a_i$.
\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurrentPolicy(a_i)}\]

\[CurrentPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The 
square root skews this probability to favor even more exploration of nodes with 
few visits.

The exploitation term is the average of the evaluation of 
all nodes descendant of $a_i$ (itself included) in the current search tree. It 
is also called the current evaluation of $a_i$.   
\[Exploitation(a_i) = CurrentEvaluation(a_i) = 
  \average_{b \in Descendant(a_i)} {PriorEvaluation(b)}\]

\subsection{Node extension}
Once a node is selected, we applied the best not-yet-applied tactics to
the open goal this node. If the tactic is productive then the we create a new 
node and calculate and save its prior evaluation. Otherwise, we create a 
non-accessible node that holds a value of 0 to notice the value function that 
node selection was a failure. This dummy node is only there to update the 
result of the Current evaluation of other nodes and is not considered in the 
calculation of exploration term and cannot be selected.

Maybe should move this defnition in static description.

@@

\begin{definition} (Productive tactic)
During a proof search, after a tactic $t$ is applied to a node with active goal 
$g$, $t$ is called 
productive if and only if all the following conditions are satisified:
\begin{itemize}
\item It does not fail. We note $l$ the list of goals returned by the tactic.
\item It does not loop. $l$ do not contain any goals that appears in its 
ancestor nodes.
\item It is not a parallel step. $l$ is not equal to or a superset of a list of 
goals produced from the same node with the same active goal.
\end{itemize}

@@

\subsection{End of the search}
The proof search stops when one of these 3 conditions is realized:
\begin{itemize}
\item It found a proof (i.e. the root node is solved). In this case, 
the search returns a minimized and 
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It has saturated. 
There is no more tactics to be applied to any open goal. \todo{define open 
goal}
\item It exceeded a given time limit.
\end{itemize}
